---
title: "p8105_hw2_zz3039.Rmd"
output: github_document
date: "2022-10-12"
---
```{r chunk_library, message=FALSE}
library(tidyverse)
library(readr)
library(p8105.datasets)
library(ggridges)
library(patchwork)
```


## Problem 2

```{r chunk_p2_1, message = FALSE}
accel_data = read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    day_type = 
      case_when(
        day %in% c("Saturday", "Sunday") ~ "weekend",
        TRUE ~ "weekday"),
    accel_day = 
      rowSums(.[4:1443]),
    day = 
      forcats::fct_relevel(day, c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")),
    day_type = 
      forcats::fct_relevel(day_type, c("weekday", "weekend"))) %>%
  arrange(week, day_id, day, day_type, accel_day) %>% 
  select(week, day_id, day, day_type, accel_day, everything())

accel_data_tidy = 
  accel_data %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "time_min",
    names_prefix = "activity_",
    values_to = "accel_minute"
  ) %>% 
  select(week, day_id, day, day_type, time_min, accel_minute, accel_day)
```
This dataset has **`r ncol(accel_data_tidy)`** variables, including **`r names(accel_data_tidy)`**, and total **`r nrow(accel_data_tidy) `** observations.

```{r chunk_p2_2}
accel_data_day =
  accel_data %>% 
  select(week, day_id, day, day_type, accel_day) 
print(accel_data_day, n = 20)
```

Actually I can't see any trends apparent here in the activity data for each day.

```{r chunk_p2_3, message = FALSE}
hour_data = 
  accel_data %>% 
  mutate(
    hour_1 = rowSums(.[6:65]),
    hour_2 = rowSums(.[66:125]),
    hour_3 = rowSums(.[126:185]),
    hour_4 = rowSums(.[186:245]),
    hour_5 = rowSums(.[246:305]),
    hour_6 = rowSums(.[306:365]),
    hour_7 = rowSums(.[366:425]),
    hour_8 = rowSums(.[426:485]),
    hour_9 = rowSums(.[486:545]),
    hour_10 = rowSums(.[546:605]),
    hour_11 = rowSums(.[606:665]),
    hour_12 = rowSums(.[666:725]),
    hour_13 = rowSums(.[726:785]),
    hour_14 = rowSums(.[786:845]),
    hour_15 = rowSums(.[846:905]),
    hour_16 = rowSums(.[906:965]),
    hour_17 = rowSums(.[966:1025]),
    hour_18 = rowSums(.[1026:1085]),
    hour_19 = rowSums(.[1086:1145]),
    hour_20 = rowSums(.[1146:1205]),
    hour_21 = rowSums(.[1206:1265]),
    hour_22 = rowSums(.[1266:1325]),
    hour_23 = rowSums(.[1326:1385]),
    hour_24 = rowSums(.[1386:1445])
  ) %>% 
  pivot_longer(
    hour_1:hour_24,
    names_to = "hour",
    names_prefix = "hour_",
    values_to = "accel_hour",
    ) %>% 
  mutate(
    hour = as.numeric(hour),
    day_id = as.character(day_id)) %>% 
  select(week, day_id, day, day_type, accel_day, hour, accel_hour)

plot_hour_data = 
  hour_data %>% 
  ggplot(aes(x = hour, y = accel_hour, fill = day_id, color = day)) +
  geom_line() +
  viridis::scale_color_viridis(
    name = "day", 
    discrete = TRUE) +
  labs(
    title = "Hourly plot of accelerometer data",
    x = "Hour of the day",
    y = "Accelerometer data") +
  scale_x_continuous(
    breaks = c(0, 12, 24),
    labels = c("0", "12", "24"),
    limits = c(0, 24)) +
  theme(legend.position = "bottom")
plot_hour_data
```

From the graph we can tell that the accelerometer data was generally low in the first 5 hours of the day, while having peaks during 5~10 hours and 20~25 hours. The weekday or weekend pattern of accelerometer data was not very clear but it can be seen that the peak in 20~25 hours was mostly caused by accelerometer data from weekdays, and the Sunday data showed unusual peak during 10~12.5 hours of the day.

## Problem 3

```{r chunk_p3_preparation}
ny_noaa %>% 
  count(id, name = "n_obs_id")
ny_noaa %>% 
  count(date, name = "n_obs_date")
na_prcp = sum(is.na(ny_noaa$prcp)) / nrow(ny_noaa)
na_prcp
na_snow = sum(is.na(ny_noaa$snow)) / nrow(ny_noaa)
na_snow
na_snwd = sum(is.na(ny_noaa$snwd)) / nrow(ny_noaa)
na_snwd
na_tmax = sum(is.na(ny_noaa$tmax)) / nrow(ny_noaa)
na_tmax
na_tmin = sum(is.na(ny_noaa$tmin)) / nrow(ny_noaa)
na_tmin
```

The dataset **ny_noaa** has **`r ncol(ny_noaa)`** variables, namely **`r names(ny_noaa)`**, with **`r nrow(ny_noaa)`** observations. The **"id"** variable includes **747** specific id numbers, and the **"date"** variable includes **10957** specific days.

The propotion of missing value in the variables **"prcp", "snow", "snwd", "tamx" and "tmin"** all exceeded **5%**, which should be paid attention to.

```{r chunk_p3_1}
noaa1 = 
  ny_noaa %>% 
  separate(date, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(
    tmax = as.numeric(tmax)/10,
    tmin = as.numeric(tmin)/10,
    month = as.integer(month),
    month = month.abb[month],
    year = as.numeric(year))
snow_data = 
  noaa1 %>% 
  count(snow, name = "n_obs", sort = TRUE)
snow_data
```
The 5 most commonly observed values in snowfall are: **0, 25, 13, 51, 76**, as can be seen from the tibble above, with null value excluded.

```{r chunk_p3_2, warning = FALSE}
jj_data = 
  noaa1 %>% 
  filter(month == "Jan"|month == "Jul")
jj_data %>% 
  group_by(year, month, id) %>% 
  mutate(mean_tmax = mean(tmax, na.rm = TRUE)) %>% 
  ggplot(aes(x = year, y = mean_tmax, color = id)) + 
  geom_line() +
  facet_grid(~month) +
  labs(
    title = "Yearly plot of max temperature",
    x = "Year",
    y = "Average max temperature"
  ) +
  theme(legend.position = "none")
```
It can be seen from the graph that the average max temperature in January and July in each station fluctuated stably across years, with temperature in January centered approximately around 0 and July around 27. There were also outliers in each month, like the extreme low temperature in July around 1982 and June around 1988.

```{r chunk_p3_3, warning = FALSE}
tmax_data = 
  noaa1 %>%
  select(year, month, tmax, tmin) %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "observation", 
    values_to = "temp") %>% 
  ggplot(aes(x = temp, fill = observation)) +
  geom_density(alpha = .5) +
  viridis::scale_fill_viridis(discrete = TRUE)

snow_data = 
  noaa1 %>% 
  filter(snow > 0 & snow < 100) %>%
  mutate(year = as.factor(year)) %>% 
  ggplot(aes(x = snow, y = year)) + 
  geom_density_ridges(scale = .85)

tmax_data/snow_data
```

The first graph showed the density of tmin and tmax, as we can see, they both followed a uniform-like distribution with the mean lower for tmin and higher for tmax.

The second graph showed the density of snowfall for each year, we can see that the heights were mostly within 0mm to 40mm.

